---
title: "Econometrics - Homework 2"
output:
  html_document: default
  html_notebook: default
  pdf_document: default
---

```{r include = FALSE}
library(stargazer)
library(dplyr)
library(GGally)
library(ggplot2)
library(skimr)
library(DataExplorer)
```

# Name: Hakon Petermann Rueda

```{r}
#loading data
prospect_data <- read.csv("prospects.csv")
credit_data <- read.csv("credit.csv")
#View(prospect_data)
#View(credit_data)

#class(credit_data$home)
#is.vector(credit_data$home)


credit_data$home_vector <- as.factor(credit_data$home)
# #class(credit_data$home_vector)
 levels(credit_data$home_vector)
# nlevels(credit_data$home_vector)
# 
credit_data$marital_vector <- as.factor(credit_data$marital)
# #class(credit_data$marital_vector)
# levels(credit_data$marital_vector)
# nlevels(credit_data$marital_vector)
# 
credit_data$job_vector <- as.factor(credit_data$job)
# #class(credit_data$job_vector)
# levels(credit_data$job_vector)
# nlevels(credit_data$job_vector)


prospect_data$home_vector <- as.factor(prospect_data$home)
prospect_data$marital_vector <- as.factor(prospect_data$marital)
prospect_data$job_vector <- as.factor(prospect_data$job)
```


Description of the variables in both data sets:
- `seniority` job seniority (years)
- `home` type of home ownership
- `age` client's age
- `marital` marital status
- `job` type of job
- `expenses` client's expenses
- `income` client's income
- `assets` client's assets
- `debt` client's existing debts
- `amount` loan amount (for prospects, this is the personal loan amount they have applied for)
- `savings` client's savings

Correlations
```{r}
#ggp <- ggpairs(credit_data[,c("default", "seniority", "home", "age", "marital", "job", "expenses", "income", "assets", "debt", "amount", "savings")])
#print(ggp, progress = FALSE)

correlations <- cor(credit_data[,c("default", "seniority", "age", "expenses", "income", "assets", "debt", "amount", "savings")])
print(correlations)
```

# 1. Build and estimate a model to predict the probability that these applicants will default on their loan (if you choose to disburse one).

## i. Write down a sensible model; explain your economic intuition for including the variables on the right-hand side of this model.

I would include most variables in the dataset. Two explanatory variables jump out as potentially being highly correlated, and this is confirmed when generating a correlation table: job *seniority* & *age*, as well as *savings level* & *income*. I will take out *age*, and *income* then because both *seniority* and *savings* have higher correlations with defaulting.

I also think that home ownership type might be correlated very highly with savings rate, and so will be removing it (i.e., those living with their parents will intuitively have less savings then home owners).

As for the rest of the numeric variables, my economic intuition says that it doesn't hurt to include them all. And for the factor variables, the same applies because there could be some hidden interactions behind the data that we do not expect. For example, marital status, or job contract type might be good predictors for default probability, but my intuition would not expect will be too highly correlated with other explanatory variables.

Estimated model:
default ~ seniority + marital + job + expenses + assets + debt + amount + savings

## ii. Estimate this model using:

### a. Linear Probability Model

```{r}
model_lm1 <- lm(default ~ seniority + marital + job_vector + expenses + assets + debt + amount + savings, data = credit_data) #as described above
model_lm2 <- lm(default ~ seniority + home + job + expenses + assets + debt + amount + savings, data = credit_data) #removed marital vector due to insignificance, added home vector instead
model_lm3 <- lm(default ~ seniority + job + expenses + assets + debt + amount + savings, data = credit_data) #removed both martial and home vector
model_lm4 <- lm(default ~ seniority + job + expenses + assets + amount + savings, data = credit_data) #removed debt due to statistical insignificance

stargazer(model_lm1, model_lm2, model_lm3, model_lm4, type = 'text')
```
Before settling on a final model, I would like to consider adding the home owner variable, as that model has the highest adjusted-R2, but I am worried that is simply due to correlation with savings rate (i.e., multicollinearity is present.). Let me use a boxplot to see this relationship visually:

```{r}
#aov1 = aov(savings ~ home_vector, data = credit_data)
#summary(aov1)
#?boxplot
boxplot(savings ~ home_vector, data = credit_data)
```

It seems that there is not a strong relationship as I expected. I was not totally wrong as the biggest and most outliers for those with the highest savings are the home owners, but those are only outliers, and they are present for the other catagories too.

Therefore my final model will be:

default ~ seniority + home + job + expenses + assets + debt + amount + savings

```{r}
#summary of chosen model
summary(model_lm2)
```


My final model was not the same as my intuitively constructed model, but it had a higher adjusted-R2 (0.185 vs 0.171).

### b. Probit

```{r}
# ?glm
model_prob1 <- glm(default ~ seniority + home + job + expenses + assets + debt + amount + savings, family = binomial(link = "probit"), data = credit_data)

summary(model_prob1)
```


### c. Logit

```{r}
# ?glm
model_log1 <- glm(default ~ seniority + home + job + expenses + assets + debt + amount + savings, family = binomial(link = "logit"), data = credit_data)

summary(model_log1)
```

### Does the model make sense? (sign and significance of variables)

Yes, the model makes sense. Increasing values for expenses, debt, loan amount, and the factors for all wokr types (besides fixed contracts) are estimated to lead to an increase in default rate. Things like job seniority, asset value, and savings amount on the other hand are estimated to have the opposite effect. As for the significance of the variables, what doesn't make sense really is that adding the home ownership type vector increases the significance of at least the debt explanatory variable, and increases the explanatory power of the model, but for the most part it is statistically insignificant.

# 2. Using the estimates from part 1.ii (for the three models separately, and for each applicant separately), predict each of these applicants’ probability of default.

```{r}
# pros1 <- data.frame(seniority = prospect_data[1,][2], home = prospect_data[1,][3], age = prospect_data[1,][4], marital = prospect_data[1,][5], job = prospect_data[1,][6], expenses = prospect_data[1,][7], income = prospect_data[1,][8], assets = prospect_data[1,][9], debt = prospect_data[1,][10], amount = prospect_data[1,][11], savings = prospect_data[1,][12])

pros <- list()

for (i in 1:10){
  pros[[i]] <- data.frame(seniority = prospect_data[i,][2], home = prospect_data[i,][3], age = prospect_data[i,][4], marital = prospect_data[i,][5], job = prospect_data[i,][6], expenses = prospect_data[i,][7], income = prospect_data[i,][8], assets = prospect_data[i,][9], debt = prospect_data[i,][10], amount = prospect_data[i,][11], savings = prospect_data[i,][12])



  cat("Prospect number", i, "(", as.character(prospect_data[i,][1]), "):")
  cat("\nLinear model probability of default:", 100*predict(model_lm2, pros[[i]]), "%")
  cat("\nProbit model probability of default:", 100*predict(model_prob1, pros[[i]], type = "response"), "%")
  cat("\nLogit model probability of default:", 100*predict(model_log1, pros[[i]], type = "response"), "%")
  cat("\n\n")

}


# predict(model_lm2, prospect_data, level=0.95, interval="confidence")
```


```{r}
# predict(model_lm2, prospect_data)
# predict(model_prob1, prospect_data, type = "response")
# predict(model_log1, prospect_data, type = "response")
```

# 3. Using the predicted default probabilities for each applicant you can now turn to making decisions on the applications. Each of these applications is for a 1-year loan, but for different amounts (see column “amount” in the prospects data). Assume that the annual interest rate that you can charge from your clients is 10%. Further assume that your cost of borrowing is zero and you have no other variable costs.

## Whom would you lend to given the predicted default probabilities and the given interest rate? Show your calculations (and try to do them in R).

Well the loan is for 1 year, with a 10% interest rate and so the profit will be 10% of the loan amount. Taking the default rates into account, we want to multiply that profit by (1 - default rate). To run through all prospects and looking at their expected profits we get (using the linear model, which in most of the 10 prospects case's is the most conservative):

```{r}

for (i in 1:10){
  expected_profit = (1-predict(model_lm2, pros[[i]])) * 0.1 * as.numeric(prospect_data[i,][11])
  cat("Prospect number", i, "(", as.character(prospect_data[i,][1]), ")", "expected profit: $", expected_profit, "\n")
}

```


## What is your expected profit if you follow this recommendation?

If we had to chose one single prospect to lend to, we would lend to the prospect number 8 (Schmidt).

It must be noted though that Schmidt's loan amount is also the largest though, so if we only had a fixed amount to loan, we would choose the prospect with the lowest default rate, which would be prospect number 9 (Oener).

##  Use the confidence interval in the linear probability model to calculate a negative case scenario: What is your expected profit if the default probabilities are at the upper value of the prediction interval.

To do this we use the predict() function for a lm and access it's third element which is the upper-bound prediction.

```{r}
for (i in 1:10){
  expected_profit = (1-predict(model_lm2, pros[[i]], level=0.95, interval="confidence")[3]) * 0.1 * as.numeric(prospect_data[i,][11])
  cat("Prospect number", i, "(", as.character(prospect_data[i,][1]), ")", "expected profit (worst-case scenario): $", expected_profit, "\n")
  
  # print(predict(model_lm2, pros[[i]], level=0.95, interval="confidence"))
  # print(predict(model_lm2, pros[[i]], level=0.95, interval="confidence")[3])
  # predict(model_lm2, pros[[i]])
}

# linear_defaults <- predict(model_lm2, prospect_data, level=0.95, interval="confidence")
# as.numeric(linear_defaults[1,][1])
```

