---
title: "Thore Sintram - Econometrics - Homework 2"
output:
  html_document: default
  html_notebook: default
---


```{r}
library(stargazer)
library(dplyr)
library(GGally)
library(ggplot2)
# rm(list =ls())
#read the data into a dataframe
credit <- read.csv(file='credit.csv', header=TRUE, sep=",")
prospects <- read.csv(file='prospects.csv', header=TRUE, sep=",")

```

Here is a description of the variables in the data set: 

- `default` 0 means they did not default, 1 means they defaulted
- `seniority` job seniority (years)
- `home` type of home ownership
- `age` client's age
- `marital` marital status
- `job` type of job
- `expenses` client's expenses
- `income` client's income
- `assets` client's assets
- `debt` client's existing debts
- `amount` loan amount (For prospects this is the personal loan amount they have applied for)
- `savings` client's savings

# Explore the data:
```{r}
View(credit)
summary(credit)
str(credit)

View(prospects)
summary(prospects)
str(prospects)

```

# Discussing a possible, sensible model
Y-variable: Default
X-variable: 
  -seniority: people with more work experience are usually more responsible
  -home: house owners are usually more capable of paying their debts
  -age: older people are usually more responsible
  -income: the better the salary, the more likely they pay back their debt (Correlation with job)
  -assets: if needed, they can pay back their debt by liquidating their assets
  -savings: the more savings, the more likely they are capable of paying back the debt
  -debt: the more debt, the less likely they pay back new debt
  -amount: the higher the amount, the less likely they are capable of paying it all back
  -expenses: the more expenses, the less likely they are gonna pay back the debt

# Estimate the model using linear probability model
```{r}
lpm1 <- lm(default ~ seniority + home + age + income + assets + savings + debt + amount + expenses, data = credit)
summary(lpm1)

```

# Estimate the model using probit
```{r}
prob1 <- glm(default ~ seniority + home + age + income + assets + savings + debt + amount + expenses, family = binomial(link="probit"), data = credit)
summary(prob1)

```

# Estimate the model using logit
```{r}
log1 <- glm(default ~ seniority + home + age + income + assets + savings + debt + amount + expenses, family = binomial(link="logit"), data = credit)
summary(log1)

```

# Do the models make sense?
There does not seem anything to be odd, the only variable, which might cause problems later on, is the low adjusted R-square of the first model.

# Predict each of of these applicants' probability of default (LPM)
```{r}
predict(lpm1, prospects, level=0.95, interval="confidence")
```
# Predict each of of these applicants' probability of default (Probit)
```{r}
predict(prob1, prospects, type = "response")
```
# Predict each of of these applicants' probability of default (Logit)
```{r}
predict(log1, prospects, type = "response")
```

# Who would you lend to given the predicted default probabilities and the given interest rate?
The closer a prospect's default raiting to 0, the more likely they are to pay back their loan.
Consequently, only prospects with a low default rating should receive a loan.

```{r}

defaulttable <- predict(lpm1, prospects, level=0.95, interval="confidence")
averagerisk <- subset(defaulttable, select=-c(lwr,upr))
profit <- select(prospects, amount) * ( 0.1 - averagerisk )
show(profit)

```
If the calculated risk is taken into account, then only person number 9 and 10 should receive a loan.
That being said, only Oener and Fassbender would receive their requested loan.

# What is your expected profit if you follow this recommendation?
```{r}
offer <- c(28.21362,11.70244)
total_profit <- sum(offer)
show(total_profit)
```

Total Profit = 329.5

# What is your expected profit if the default probabilities are at the upper value of the prediction interval?

Generate a data frame with data for the worst case scenario

```{r}
worstcase <- subset(defaulttable, select=-c(lwr,fit))
worstprofit <- select(prospects, amount) * ( 0.1 - worstcase )
show(worstprofit)
```

In the worst case scenario, only person number 9 (Oener) would receive the requested loan.
The expected total profit would therefore be ≈ 9.744
